Coalescent inference of mutation spectrum histories from sample frequency spectra

### 2019-08-26 manuscript/main.tex
he's into it
a schematic figure of our model

\usepackage{fullpage}
\usepackage[round,semicolon]{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}

\usepackage{hyperref}

@@

\author{
W. DeWitt$^{1}$, K. Harris (if he's into it?)$^{2}$, and K. Harris$^{1}$\\
\small{Departments of $^1$Genome Sciences and $^2$Biology, University of Washington, Seattle, USA}
% \small{$^\ast$ Equal contribution}
}

@@

The setting for our modeling is Kingman's coalescent \citep{Kingman1982-ge, Kingman1982-tf, Kingman1982-ys, Kingman2000-jr}, with all the usual niceties: neutrality, infinite sites, linkage equilibrium, and panmixia.
We will retrace the derivation by \cite{Griffiths1998-qf} of the frequency distribution of a derived allele conditioned on the demographic history, while generalizing to a time inhomogeneous mutation process.
We will make use of important results of \cite{Polanski2003-kg} and \cite{Polanski2003-ll} which facilitate numerical computation under this model, and formulate the expected 3-SFS under given demographic and mutation spectrum histories following the notation of \cite{Rosen2018-bb} for the expected SFS with a constant mutation process.

@@

@@

Todo:
\begin{itemize}
\item repeat 1KG analysis like in \cite{Harris2017-fw}, see if we recapitulate Kelley's simulation-based pulse results
\item cluster triplet time series to see if we pull in minor components. Dynamic time warping?
\end{itemize}

>>>>

\usepackage{fullpage}
\usepackage[round,semicolon]{natbib}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bbm}
\usepackage{graphicx}
\usepackage{hyperref}

@@

\author{
W. DeWitt$^{1}$, K. Harris$^{2}$, and K. Harris$^{1}$\\
\small{Departments of $^1$Genome Sciences and $^2$Biology, University of Washington, Seattle, USA}
% \small{$^\ast$ Equal contribution}
}

@@

The setting for our modeling is Kingman's coalescent \citep{Kingman1982-ge, Kingman1982-tf, Kingman1982-ys, Kingman2000-jr}, with all the usual niceties: neutrality, infinite sites, linkage equilibrium, and panmixia.
We will retrace the derivation by \cite{Griffiths1998-qf} of the frequency distribution of a derived allele conditioned on the demographic history, while generalizing to a time inhomogeneous mutation process.
We will make use of important results of \cite{Polanski2003-kg} and \cite{Polanski2003-ll} which facilitate numerical computation under this model, and formulate the expected 3-SFS under piecewise constant demographic and mutation spectrum histories building on the notation of \cite{Rosen2018-bb} for the expected SFS with a piecewise constant demographic history and constant mutation process.

@@

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/model}
\caption{Schematic of the marked Poisson process model with $n=4$.
We condition on coalescent times $T_4=t_4,T_3=t_3,T_2=t_2$ and consider mutation intensity function $\mu(t)$.
Red dots indicate mutation events placed by time $t$, genomic position $x$, and coalescent line (which are depicted as extruded in the genomic coordinate axis, grey rectangles).
The probability that a differential element $dxdt$ on a given sheet contains a mutation is proportional to the instantaneous mutation intensity $\mu(t)$.
}
\label{fig:model}
\end{figure}

@@

Todo:
\begin{itemize}
\item repeat 1KG analysis like in \cite{Harris2017-fw}, see if we recapitulate Kelley's simulation-based pulse results
\item cluster triplet time series to see if we pull in minor components.
\end{itemize}
